---
title: "p8105_hw2_cl4043"
author: Stella Li
output: github_document
---
This is my solution for hw2.

```{r message=FALSE}
library(tidyverse)
library(readxl)
library(readr)
```

## Problem 1

```{r}
# First, define a path to the dataset. 
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```


```{r}
# Read the Mr. Trashwheel dataset.
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```


```{r}
# Read precipitation data! For 2018 and 2017. 
precip_2018 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
precip_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation dataframes. In the following code chunk, I create a "helper" tibble that contains pairs of numeric and character ways of representing month, and then merge that (using month number as a key) with the precipitation dataset. This technique is one I use often when I need to recode a moderate or large number of values for a variable. 

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)
precip_df =
	left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2

```{r}
# define a path to the dataset. 
path_to_data2 = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"
```

```{r message=FALSE}
# read the nyc transit data
# retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance
nyc_transit =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"
           ) %>%
  janitor::clean_names() %>%
  select(line:route11, entrance_type, entry, vending, ada) %>%
  mutate_at(vars(route1:route11), as.factor) %>%
  mutate(entry = recode(entry, `YES` = TRUE, `NO` = FALSE)) # converting entry variable to a logical variable
```

```{r}
# tidying nyc_transit data to a long table by route number
nyc_transit_tidy = 
  pivot_longer(
    nyc_transit, 
    route1:route11,
    names_to = "route_number", 
    values_to = "route")

# dropping the NA value when route value contains NA of each route number
nyc_transit_tidy <- nyc_transit_tidy %>% 
  drop_na()
```

This dataset contains information about the train transit system in NYC. The dataset contains variables such as, `r names(nyc_transit_tidy)`. I first used the cleaned the names of each variable, and then I used the select function to only retain the variables of interest. I noticed there were both characters and numerics in the values of route number, so I turned them all into factor. I realized the route numbers columns were spread across the original dataset, so I used pivot_longer to make it into a long table, and then dropped the rows where the route value is NA. The dimension of the resulting dataset is `r dim(nyc_transit_tidy)`, which means there are in total 4270 rows of observations, and 10 columns.

* There are `r nrow(distinct(nyc_transit_tidy, station_name, line))` distinct stations.
```{r, echo=FALSE}
distinct_stations<- distinct(nyc_transit_tidy, station_name, line, ada, vending)
```
* Among the distinct stations, there are `r nrow(filter(distinct_stations, ada == "TRUE"))` stations that are ADA compliant.
* There are `r nrow(filter(nyc_transit_tidy,route ==  "A"))` distinct stations that serve the A station, and among these stations, `r nrow(filter(nyc_transit_tidy,route == "A" & ada == "TRUE"))` stations are ADA compliant.
* The proportion of station entrances / exits without vending allow entrance is `r nrow(filter(nyc_transit_tidy,vending == "NO"))/nrow(nyc_transit_tidy)`.

## Problem 3

```{r}
# define paths to the datasets. 
path_to_data3 = "./data/pols-month.csv"
path_to_data4 = "./data/snp.csv"
path_to_data5 = "./unemployment.csv"
```

```{r message=FALSE}
# read the pols-month.csv
# separate() to break up the variable mon into integer variables year, month, and day
pols_month_df =
  read_csv("./data/pols-month.csv"
           ) %>%
  janitor::clean_names() %>%
  separate(mon, c("year", "month", "day")) %>%
  mutate_at(vars(month), as.factor)

# replace month number with month name
pols_month_df$month <- month.abb[pols_month_df$month]

# create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.
pols_month_df <- pols_month_df %>%
  mutate(president = case_when(
    prez_gop == 1 ~ 'gop',
    prez_dem == 1 ~ 'dem'
)) 
pols_month_df <- pols_month_df[-c(3,4,8)]
```


```{r, message=FALSE}
# separate() to break up the variable mon into integer variables year, month, and day
# arrange according to year and month, and organize so that year and month are the leading columns
snp_df =
  read_csv("./data/snp.csv"
           ) %>%
  janitor::clean_names() %>%
  separate(date, c("day", "month", "year")) %>%
  arrange(year, month) %>%
  relocate(year, month) %>%
  mutate_at(vars(month), as.factor)

snp_df$month <- month.abb[snp_df$month]
```

```{r, message=FALSE}
unemployment_df =
  read_csv("./data/unemployment.csv"
           ) 
  

unemployment_long = 
  pivot_longer(
    unemployment_df, 
    Jan:Dec,
    names_to = "month", 
    values_to = "unemployment_rate")

merge_df = 
  left_join(pols_month_df, snp_df, by = "month")

merge_final = 
  left_join(merge_df, unemployment_long, by ="month")
```

The dimension of the resulting dataset is `r dim(merge_final)`, and contains variables such as "month", "unemployment_rate", and "president". The range of the years is `r range(merge_final$Year)`.
